---
title: "Basic Plots in R"
output: html_document
---


##### Scatterplot


Basic command plot maps a scatterplot in R. Setting type to **type="l"** creates a line plot.

To add additional plots on top of the original plot, use functions **lines()** to add lines and function **points()** to add scaterplots.


```{r, message=FALSE, warning=FALSE, cache=TRUE}
data(wage1, package='wooldridge')
plot(wage1$educ, wage1$wage,
    main="Education and Wages",
    ylab="Wage",
    xlab="Education",
    type="p",
    col="blue")

lines( seq(min(wage1$educ),max(wage1$educ),1) , rep(mean(wage1$wage),range(wage1$educ)[2]+1), col="red")

legend("topleft",
c("scater","mean wage"),
fill=c("blue","red")
)
```

##### Scatterplot with Subgroups

```{r, message=FALSE, warning=FALSE, cache=TRUE}
#Separating data into two groups
wage2=wage1[which(wage1$wage>=7.5),c(1,2)]
wage3=wage1[which(wage1$wage<7.5),c(1,2)]
#Using plot to plots the first group and points function for the second group 
plot(wage~educ, data=wage3, pch=1, col="blue", 
     ylim=c(0,26), xlim=c(0,20), main="Wages and Education")
points(wage~educ, data=wage2, pch=2, col="green")
```

##### Simple Barplot

Function **barplot** displays a simple bar plot.

```{r, message=FALSE, warning=FALSE, cache=TRUE}
par(mfrow=c(1,2))
barplot(sort(wage1$wage))
barplot(sort(wage1$wage[1:25]))
```

##### Pie

```{r, message=FALSE, warning=FALSE, cache=TRUE}
pie(c(sum(wage1$female),sum(!wage1$female)),
    labels=c("Female","Male"))
```

##### Boxplot

```{r, message=FALSE, warning=FALSE, cache=TRUE}
boxplot(wage1$wage,wage1$educ,
        main="Boxplot for Wages and Education",
        xlab=c("Wages and Education") )
```

```{r, message=FALSE, warning=FALSE, cache=TRUE}
boxplot(wage1$exper[wage1$female==1],wage1$exper[wage1$female==0],
        main="Boxplot for Wages and Education",
        xlab=c("Wages and Education"),
        col=c("gold","darkgreen"),
        notch=TRUE)
```

```{r, message=FALSE, warning=FALSE, cache=TRUE}
#install.packages("vioplot")
library(vioplot)
vioplot(wage1$exper[wage1$female==1], wage1$exper[wage1$female==0], names=c("Female", "Male"),
   col="gold",
   main="Experience by Gender")
```

##### Histogram and Density

```{r, message=FALSE, warning=FALSE, cache=TRUE}
hist(wage1$wage, breaks=10, main="Histogram of Wages", ylim=c(0,210), labels=TRUE)
```

##### Histogram and Density

```{r, message=FALSE, warning=FALSE, cache=TRUE}
hist(log(wage1$wage), freq=FALSE,ylim=c(0,1),
     main="Histogram")
lines(density(log(wage1$wage)),col=4)
```

##### Plotting OLS Regression Line (Simple Regression)

```{r, message=FALSE, warning=FALSE, cache=TRUE}
reg1=lm(log(wage)~educ, data=wage1)
plot(wage~educ, data=wage1)
a0=1/length(fitted(reg1)) * sum(exp(reg1$residuals))
points(exp(reg1$fitted.values)*a0~wage1$educ,lwd=10,col=2)
```

##### Plotting OLS Regression Line (Multiple Regression)

```{r, message=FALSE, warning=FALSE, cache=TRUE}
reg2=lm(log(wage)~educ+exper,data=wage1)
plot(wage~exper,data=wage1)
a1=1/length(fitted(reg1)) * sum(exp(reg1$residuals))
yfit=reg2$coefficients[1]+reg2$coefficients[2]*mean(wage1$educ)+reg2$coefficients[3]*(min(wage1$exper):max(wage1$exper))
points(min(wage1$exper):max(wage1$exper),exp(yfit)*a1,lwd=5,col=6)
```

##### Plotting OLS Regression Line with Confidence Intervals

```{r, message=FALSE, warning=FALSE, cache=TRUE}
reg3=lm(log(wage)~educ, data=wage1)
reg3predict=predict(reg3, interval="prediction", newdata=data.frame(educ=seq(from=0, to=18, by=1)))
plot(wage~educ, data=wage1)
a2=1/length(fitted(reg3)) * sum(exp(reg3$residuals))
cor_reg3predict=a2*exp(reg3predict)
lines(cor_reg3predict[,1]~seq(from=0, to=18, by=1), col=1)
lines(cor_reg3predict[,2]~seq(from=0, to=18, by=1), col=1, lty=2)
lines(cor_reg3predict[,3]~seq(from=0, to=18, by=1), col=1, lty=2)
```


##### Diagnostic Plots and Box-Cox Transformation

```{r 1, message=FALSE, warning=FALSE, cache=TRUE}
library(car)
reg4=lm(wage~educ, data=wage1)
par(mfrow=c(2,2))
plot(wage~educ, data=wage1)
hist(reg4$residuals)
qqPlot(rstandard(reg4))
plot(wage~educ, data=wage1)
points(reg4$fitted.values~wage1$educ,lwd=5,col=2)
#Box-Cox Transformation
bc1=boxCox(wage1$wage ~ wage1$educ, lambda = seq(-0.5, 0.5, length = 10))
lambda=bc1$x
likel=bc1$y
lamlike=cbind(lambda,likel)
sorted=lamlike[order(-likel),]
head(sorted,n=5)
reg4b=lm(wage^(sorted[1])~educ, data=wage1)
plot(wage~educ, data=wage1)
hist(reg4b$residuals)
qqPlot(rstandard(reg4b))
#plot(wage~educ, data=wage1)
#points((reg4b$fitted.values)^(1/sorted[1])~wage1$educ,lwd=5,col=2)
#Compare before and after
par(mfrow=c(3,2))
plot(wage~educ, data=wage1)
points(reg4$fitted.values~wage1$educ,lwd=5,col=2)
plot(wage~educ, data=wage1)
points((reg4b$fitted.values)^(1/sorted[1])~wage1$educ,lwd=5,col=2)
hist(reg4$residuals)
hist(reg4b$residuals)
qqPlot(rstandard(reg4))
qqPlot(rstandard(reg4b))
```


---
